---
title: "Introduction to fibga: Simulating and Evolving Networks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to fibga: Simulating and Evolving Networks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(fibga)
```

# Overview

The `fibga` package implements a genetic algorithm for discovering parameter sets
that encourage amyloid fibril formation in network models. It simulates networks using ERGM models,
selects high-yield topologies, and evolves parameters across generations.

This vignette walks through a basic example of using the package.

# Basic Concepts

- **Breeding:** Create offspring parameter sets by interpolating between parent vectors.
- **Mutation:** Add random noise to parameter sets.
- **Evolution:** Select the best-performing parameter sets and evolve them across generations.
- **Fibril Assay:** Classify network nodes into fibril structures based on orbit counts.

# Quick Example: Running Full Evolution

Let's run a very small example:

```{r full-evolution-example, eval=TRUE}
set.seed(0)
# Define an initial parameter set
init_param <- c(100, -30, 0.5, 5)

# Run the evolution pipeline
results <- run_full_evolution(init_param = init_param,
                              spheroid_pt_count = 40,
                              radius = 2,
                              node_count = 48,
                              suff_stats = "edges+kstar(2)+nsp(1:2)",
                              fib_type = "2-ribbon",
                              num_generations = 3,
                              burn_time = 2^18,
                              np = 1)

best_params <- results$final_sims[[which.max(results$mean_fibril_fractions)]]$params[[1]]
# Plot mean fibril yields across generations
plot(results$mean_fibril_fractions, type = "b",
     main = "Mean Fibril Yields Across Evolution",
     ylab = "Mean Fibril Yield")
```

# Understanding the Output

- `results$generations` contains the parameter sets that survived each generation.
- `results$final_sims` contains simulations run with the final parameters.
- `results$mean_fibril_fractions` shows how the mean fibril yield improves over generations.

# Advanced Options

You can adjust several settings in `run_full_evolution`:

- **`burn_time`:** Increase for more thorough ERGM sampling.
- **`generations`:** Evolve over more generations.
- **`spheroid_pt_count`:** Sample more initial parameter points.
- **`np`:** Increase number of CPU cores for faster simulation.

# Notes and References

- Fibril topology classification uses orbit counts computed via the ORCA software.
- For more details, see:
  - ORCA documentation: [JSS article](https://www.jstatsoft.org/article/view/v071i10/0)
  - ergm.graphlets package: [CRAN page](https://CRAN.R-project.org/package=ergm.graphlets)
---